Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes

### 2020-01-14 ms/SI.rmd
Change location of ms output to results/
Add docx conversion for SI

---
title: ""
bibliography: "si_references.bib"
mainfont: Times New Roman.ttf
fontsize: 12pt
mainfontoptions: 
- BoldFont=Times New Roman Bold.ttf
- ItalicFont=Times New Roman Italic.ttf
- BoldItalicFont=Times New Roman Bold Italic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "new-phytologist.csl"]
    includes:
      in_header: newphyt.sty
nocite: | 
  @NMNS2008, @Copeland1932, @FloraNSW2019, @Brownsey1987, @Murdock2003, @Nayar1971, @Lloyd1980, @Zhang2008, @Bierhorst1967, @Martin2006, @Tigerschiold1989, @Tigerschiold1990, @Atkinson1975, @Palmer2003
---

@@

Article title: Constrained, yet diverse: Phylogenetic and morphological filtering in epiphytic ferns

@@

**Table S1** List of ferns from Moorea, French Polynesia included in this study.

@@

**Table S2** Trait values of ferns on Moorea, French Polynesia.

SLA, specific leaf area.

Continuous measurements shown as mean, standard deviation (sample size). 'taxon code' is the name used for that taxon in analyses (see Table S1). For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI.

```{r}
trait_data_for_si %>%
  write_csv(here::here("ms/Table_S2.csv"))
```

**Table S3** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate.

@@

```{r climate-ancova}
# Format and write out table as CSV file
climate_ancova_results %>% 
  mutate(
    p.value = signif(p.value, 2) %>% format.pval(., digits = 1),
    meansq = round_t(meansq, 2),
    term = case_when(
      term == "habit" ~ "Growth habit",
      term == "el" ~ "Elevation",
      term == "habit:el" ~ "Growth habit × elevation",
      TRUE ~ term
    )
  ) %>%
  select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value) %>%
  write_csv(here::here("ms/Table_S3.csv"))

@@

**Table S4** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.

P_rnd, probability of obtaining the observed value of *D* under random trait evolution; P_bm, probability of obtaining the observed value of *D* under simulation of trait evolution by Brownian motion.

```{r phylosig-binary-strict}
# Format and write out table as CSV file
phylosig_binary_traits_strict %>%
  mutate(D = round_t(D, 2)) %>%
  mutate(trait = factor(
    trait, 
    levels = c("habit", "gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Epiphytic growth", "Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
  arrange(trait) %>%
  select(Trait = trait, `Number of presences` = num_present, `Number of absences` = num_absent, D,
         P_rnd = prob_random, P_bm = prob_brownian) %>%
  write_csv(here::here("ms/Table_S4.csv"))
```

**Table S5** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy.

LL, log likelihood.

Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs. terrestrial). Lower log likelihood for the dependent model indicates correlation with growth habit.

```{r corr-evo-strict}
# Format and write out table as CSV file
correlated_evo_test_strict %>%
  mutate(trait = factor(
    trait, 
    levels = c("gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Gemmae", "Glands", "Hairs", "Morphotype"))) %>%
  arrange(trait) %>%
  select(
    Trait = trait,
    `LL (independent model)` = logL_indep,
    `LL (dependent model)` = logL_dep,
    `Likelihood ratio` = likelihood_ratio,
    `P` = pval
  ) %>%
  write_csv(here::here("ms/Table_S5.csv"))
```

**Table S6** Best-fitting models explaining community diversity metrics by climate and growth habit.

MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, *P* value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's *I*.

Models within delta AICc 3 of the best-fitting model shown. Trait names refer to community-weighted mean values (see Methods). Frond length and width were highly correlated with rhizome diameter and not shown. For details on calculations of MPD and MNTD, see Methods.

```{r best-fit-mods}
# Format and write out table as CSV file
best_fit_div_models_moran %>%
  # Drop frond length and width (highly correlated with rhizome dia.)
  filter(resp_var %in% c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna")) %>%
  mutate(resp_var = factor(
        resp_var, 
        levels = c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna"))) %>%
  mutate(
    resp_var = lvls_revalue(
        resp_var, 
        new_levels = c("Richness", "MPDphy", "MNTDphy", 
                       "MPDfunc", "MNTDfunc",
                       "Dissection", "SLA", "Stipe length", "Rhizome diam.", "Pinna no.")) 
  ) %>%
  arrange(resp_var, delta_AICc) %>%
  mutate(
    modname = modname %>%
      str_replace_all("\\+", " \\+ ") %>%
      str_replace_all("\\.by\\.", " x ") %>%
      str_replace_all("habit", "Growth habit") %>%
      str_replace_all("temp_mean", "Mean temp.") %>%
      str_replace_all("temp_sd", "SD temp.") %>%
      str_replace_all("temp_max", "Max. temp.") %>%
      str_replace_all("vpd_mean", "Mean VPD") %>%
      str_replace_all("vpd_sd", "SD VPD") %>%
      str_replace_all("vpd_min", "Min. VPD")
  ) %>% 
  select(
    `Metric` = resp_var,
    `Model` = modname,
    AICc,
    `R2` = r2,
    delta_AICc,
    `Moran's I` = morans_I,
    `P_moran` = morans_I_pval
  ) %>%
  write_csv(here::here("ms/Table_S6.csv"))
```

**Fig. S1** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input.

**(a)**, standard effect size of mean phylogenetic distance (MPD~func~) **(b)**, standard effect size of mean nearest taxon distance (MNTD~func~). For details on calculations of MPD~func~ and MNTD~func~, see Methods. Correlations (values in each tile) calculated with Pearson's correlation coefficient using the R package "corrr". All trait sets were scaled to standard deviation of each trait. The following trait sets were used: "all" (all traits), "size" (sporophyte traits related to size; frond length and width, stipe length, rhizome width, pinna number, SLA, degree of lamina dissection), "other" (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), "gameto" (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), "reduced" (all traits except for frond length and width, which were highly correlated with stipe length and rhizome diameter). For trait sets that included traits with skewed distributions in their raw values (frond length and width, stipe length, and rhizome width), both log-transformed and untransformed sets were included. The log-transformed "reduced" dataset was used for the main results.

@@

ggsave(plot = fig_s1, filename = here::here("ms/Fig_S1.eps"), height = 8, width = 4, units = "in")
```

**Fig. S2** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia. Epiphytic communities in green, terrestrial communities in brown. Points randomly jittered along x-axis to reduce overlap. Boxplots show median values (bold lines) by growth habit. Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 $\times$ the interquartile range. Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; $*** = \pval<0.001$; $** = \pval<0.01$; $* = \pval<0.05$.

**(a)**, traits related to plant size **(b)**, traits not related to plant size.

```{r fig-s1}
loadd(cwsd_scatterplot, cache = canopy_cache)
ggsave(plot = cwsd_scatterplot, filename = here::here("ms/Fig_S2.eps"), height = 6, width = 7, units = "in")
```

>>>>

---
title: ""
bibliography: "si_references.bib"
mainfont: Times New Roman.ttf
fontsize: 12pt
mainfontoptions: 
- BoldFont=Times New Roman Bold.ttf
- ItalicFont=Times New Roman Italic.ttf
- BoldItalicFont=Times New Roman Bold Italic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    pandoc_args: [ "--csl", "new-phytologist.csl"]
    includes:
      in_header: new-phytologist.sty
nocite: | 
  @NMNS2008, @Copeland1932, @FloraNSW2019, @Brownsey1987, @Murdock2003, @Nayar1971, @Lloyd1980, @Zhang2008, @Bierhorst1967, @Martin2006, @Tigerschiold1989, @Tigerschiold1990, @Atkinson1975, @Palmer2003
---

@@

Article title: Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes

@@

**Table S1** List of ferns from Moorea, French Polynesia included in this study

@@

**Table S2** Trait values of ferns on Moorea, French Polynesia

SLA, specific leaf area. Continuous measurements shown as mean, standard deviation (sample size). 'taxon code' is the name used for that taxon in analyses (see Table S1). For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI.

```{r}
trait_data_for_si %>%
  write_csv(here::here("results/Table_S2.csv"))
```

**Table S3** Analysis of covariance (ANCOVA) for climate variables in response to growth habit (epiphytic vs. terrestrial) with elevation as a covariate

@@

```{r climate-ancova}
# Format and write out table as CSV file
climate_ancova_results %>% 
  mutate(
    p.value = signif(p.value, 2) %>% format.pval(., digits = 1),
    meansq = round_t(meansq, 2),
    term = case_when(
      term == "habit" ~ "Growth habit",
      term == "el" ~ "Elevation",
      term == "habit:el" ~ "Growth habit × elevation",
      TRUE ~ term
    )
  ) %>%
  select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value) %>%
  write_csv(here::here("results/Table_S3.csv"))

@@

**Table S4** Phylogenetic signal in binary traits, strict dataset excluding any species whose traits were scored based on taxonomy

P_rnd, probability of obtaining the observed value of \D under random trait evolution; P_bm, probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.

```{r phylosig-binary-strict}
# Format and write out table as CSV file
phylosig_binary_traits_strict %>%
  mutate(D = round_t(D, 2)) %>%
  mutate(trait = factor(
    trait, 
    levels = c("habit", "gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Epiphytic growth", "Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
  arrange(trait) %>%
  select(Trait = trait, `Number of presences` = num_present, `Number of absences` = num_absent, D,
         P_rnd = prob_random, P_bm = prob_brownian) %>%
  write_csv(here::here("results/Table_S4.csv"))
```

**Table S5** Pagel's (1994) test of correlated evolution between binary traits, strict dataset excluding any species whose traits were scored based on taxonomy

LL, log likelihood. Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs. terrestrial). Lower log likelihood for the dependent model indicates correlation with growth habit.

```{r corr-evo-strict}
# Format and write out table as CSV file
correlated_evo_test_strict %>%
  mutate(trait = factor(
    trait, 
    levels = c("gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Gemmae", "Glands", "Hairs", "Morphotype"))) %>%
  arrange(trait) %>%
  select(
    Trait = trait,
    `LL (independent model)` = logL_indep,
    `LL (dependent model)` = logL_dep,
    `Likelihood ratio` = likelihood_ratio,
    `P` = pval
  ) %>%
  write_csv(here::here("results/Table_S5.csv"))
```

**Table S6** Best-fitting models explaining community diversity metrics by climate and growth habit

MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, *P* value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's *I*. Models within delta AICc 3 of the best-fitting model shown. Trait names refer to community-weighted mean values (see Methods). Frond length and width were highly correlated with rhizome diameter and not shown. For details on calculations of MPD and MNTD, see Methods.

```{r best-fit-mods}
# Format and write out table as CSV file
best_fit_div_models_moran %>%
  # Drop frond length and width (highly correlated with rhizome dia.)
  filter(resp_var %in% c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna")) %>%
  mutate(resp_var = factor(
        resp_var, 
        levels = c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna"))) %>%
  mutate(
    resp_var = lvls_revalue(
        resp_var, 
        new_levels = c("Richness", "MPDphy", "MNTDphy", 
                       "MPDfunc", "MNTDfunc",
                       "Dissection", "SLA", "Stipe length", "Rhizome diam.", "Pinna no.")) 
  ) %>%
  arrange(resp_var, delta_AICc) %>%
  mutate(
    modname = modname %>%
      str_replace_all("\\+", " \\+ ") %>%
      str_replace_all("\\.by\\.", " x ") %>%
      str_replace_all("habit", "Growth habit") %>%
      str_replace_all("temp_mean", "Mean temp.") %>%
      str_replace_all("temp_sd", "SD temp.") %>%
      str_replace_all("temp_max", "Max. temp.") %>%
      str_replace_all("vpd_mean", "Mean VPD") %>%
      str_replace_all("vpd_sd", "SD VPD") %>%
      str_replace_all("vpd_min", "Min. VPD")
  ) %>% 
  select(
    `Metric` = resp_var,
    `Model` = modname,
    AICc,
    `R2` = r2,
    delta_AICc,
    `Moran's I` = morans_I,
    `P_moran` = morans_I_pval
  ) %>%
  write_csv(here::here("results/Table_S6.csv"))
```

**Fig. S1** Correlation plot comparing results of functional community diversity analyses using different sets of traits as input. **(a)**, standard effect size of mean phylogenetic distance (MPD~func~) **(b)**, standard effect size of mean nearest taxon distance (MNTD~func~). For details on calculations of MPD~func~ and MNTD~func~, see Methods. Correlations (values in each tile) calculated with Pearson's correlation coefficient using the R package 'corrr'. All trait sets were scaled to standard deviation of each trait. The following trait sets were used: 'all' (all traits), 'size' (sporophyte traits related to size; frond length and width, stipe length, rhizome width, pinna number, SLA, degree of lamina dissection), 'other' (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), 'gameto' (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), 'reduced' (all traits except for frond length and width, which were highly correlated with stipe length and rhizome diameter). For trait sets that included traits with skewed distributions in their raw values (frond length and width, stipe length, and rhizome width), both log-transformed and untransformed sets were included. The log-transformed 'reduced' dataset was used for the main results.

@@

ggsave(plot = fig_s1, filename = here::here("results/Fig_S1.eps"), height = 8, width = 4, units = "in")
```

**Fig. S2** Comparison of community-weighted standard deviations (CWSD) in traits of epiphytic and terrestrial fern communities on Moorea, French Polynesia. Epiphytic communities in green, terrestrial communities in brown. Points randomly jittered along x-axis to reduce overlap. Boxplots show median values (bold lines) by growth habit. Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range. Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \*\* = \pval{ }< 0.01; \* = \pval{ }< 0.05 **(a)**, traits related to plant size **(b)**, traits not related to plant size.

```{r fig-s1}
loadd(cwsd_scatterplot, cache = canopy_cache)
ggsave(plot = cwsd_scatterplot, filename = here::here("results/Fig_S2.eps"), height = 6, width = 7, units = "in")
```

### 2020-01-14 ms/manuscript.Rmd
Change location of ms output to results/
Add docx conversion for SI

---
title: "Life in the canopy: community trait assessments reveal substantial functional diversity among fern epiphytes"
bibliography: "references.bib"
mainfont: Times New Roman.ttf
fontsize: 12pt
mainfontoptions: 
- BoldFont=Times New Roman Bold.ttf
- ItalicFont=Times New Roman Italic.ttf
- BoldItalicFont=Times New Roman Bold Italic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes # need for converting tex to docx
    latex_engine: xelatex
    pandoc_args: [ "--csl", "new-phytologist.csl"]
    includes:
      in_header: new-phytologist.sty
header-includes: 
  - \usepackage{float}
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # Fix position of figs
  - \makeatletter\renewcommand*{\fps@table}{H}\makeatother # Fix position of tables
  - \usepackage{caption} # Needed to turn off automatic numbering of table captions
editor_options: 
  chunk_output_type: console
nocite: | 
  @Vogel1968
---

@@

ggsave(plot = climate_plot, filename = here::here("ms/Fig_2.eps"), height = 4, width = 7, units = "in", device = cairo_ps_high_res)
```

@@

```{r pca, fig.height = 6, fig.width = 6}
pca_plot
ggsave(plot = pca_plot, filename = here::here("ms/Fig_3.eps"), height = 6, width = 6, units = "in", device = cairo_ps_high_res)
```

@@

```{r tree, fig.height = 8, fig.width = 6}
traits_with_tree
ggsave(plot = traits_with_tree, filename = here::here("ms/Fig_4.eps"), height = 8, width = 6, units = "in", device = cairo_ps_high_res)
```

@@

```{r heatmap, fig.height = 4, fig.width = 6}
importance_heatmap
ggsave(plot = importance_heatmap, filename = here::here("ms/Fig_5.eps"), height = 4, width = 6, units = "in", device = cairo_ps_high_res)
```

@@

```{r cwm-div, fig.height = 4, fig.width = 7}
combined_cwm_plots
ggsave(plot = combined_cwm_plots, filename = here::here("ms/Fig_6.eps"), height = 4, width = 7, units = "in", device = cairo_ps_high_res)
```

@@

```{r comm-div, fig.height = 8, fig.width = 4}
combined_comm_div_plots
ggsave(plot = combined_comm_div_plots, filename = here::here("ms/Fig_7.eps"), height = 8, width = 4, units = "in", device = cairo_ps_high_res)
```

>>>>

---
title: "Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes"
bibliography: "references.bib"
mainfont: Times New Roman.ttf
fontsize: 12pt
mainfontoptions: 
- BoldFont=Times New Roman Bold.ttf
- ItalicFont=Times New Roman Italic.ttf
- BoldItalicFont=Times New Roman Bold Italic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes # need for converting tex to docx
    latex_engine: xelatex
    pandoc_args: [ "--csl", "new-phytologist.csl"]
    includes:
      in_header: new-phytologist.sty
header-includes: 
  - \usepackage{float}
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # Fix position of figs
  - \makeatletter\renewcommand*{\fps@table}{H}\makeatother # Fix position of tables
  - \usepackage{caption} # Needed to turn off automatic numbering of table captions
editor_options: 
  chunk_output_type: console
nocite: | 
  @Vogel1968
---

@@

ggsave(plot = climate_plot, filename = here::here("results/Fig_2.eps"), height = 4, width = 7, units = "in", device = cairo_ps_high_res)
```

@@

```{r pca, fig.height = 6, fig.width = 6}
pca_plot
ggsave(plot = pca_plot, filename = here::here("results/Fig_3.eps"), height = 6, width = 6, units = "in", device = cairo_ps_high_res)
```

@@

```{r tree, fig.height = 8, fig.width = 6}
traits_with_tree
ggsave(plot = traits_with_tree, filename = here::here("results/Fig_4.eps"), height = 8, width = 6, units = "in", device = cairo_ps_high_res)
```

@@

```{r heatmap, fig.height = 4, fig.width = 6}
importance_heatmap
ggsave(plot = importance_heatmap, filename = here::here("results/Fig_5.eps"), height = 4, width = 6, units = "in", device = cairo_ps_high_res)
```

@@

```{r cwm-div, fig.height = 4, fig.width = 7}
combined_cwm_plots
ggsave(plot = combined_cwm_plots, filename = here::here("results/Fig_6.eps"), height = 4, width = 7, units = "in", device = cairo_ps_high_res)
```

@@

```{r comm-div, fig.height = 8, fig.width = 4}
combined_comm_div_plots
ggsave(plot = combined_comm_div_plots, filename = here::here("results/Fig_7.eps"), height = 8, width = 4, units = "in", device = cairo_ps_high_res)
```
